Population structure and phenotypic variation of \emph{Sclerotinia sclerotiorum} from dry bean (\emph{Phaseolus vulgaris}) in the United States

### 2017-11-17 doc/manuscript/manuscript_shell.tex
run peerj-nitpick.sh

\section*{Introduction}\label{introduction}
\addcontentsline{toc}{section}{Introduction}

\emph{Sclerotinia sclerotiorum} (Lib.) de Bary is an ascomycete plant
pathogen with a worldwide distribution (Bolton et al., 2006). This is a
necrotrophic pathogen that is primarily homothallic (self-fertilization)
and has the ability to survive for more than five years in soil using
melanized survival structures called sclerotia (Bolton et al., 2006;
Sexton et al., 2006). It causes disease on more than 400 plant species
belonging to 75 families (Boland \& Hall, 1994) including crops of major
economic importance such as sunflower (\emph{Helianthus spp.}), soybean
(\emph{Glycine max} L.), canola (\emph{Brassica napa} L., \emph{Brassica
campestris} L.), and dry bean (\emph{Phaseolus vulgaris} L.) (Bolton et
al., 2006).

On dry bean, \emph{S. sclerotiorum} is the causal agent of white mold, a
devastating disease that can be yield-limiting in temperate climates
(Steadman, 1983). All above-ground tissues (flowers, stems, leaves,
pods) are susceptible to infection, first appearing as wet lesions with
white mycelial tufts, and then bleaching as the tissue senesces
(Steadman, 1983; Bolton et al., 2006). For many years, white mold has
been the most serious dry bean disease in the Northwestern United States
(Otto-Hanson et al., 2011; Knodel et al., 2012, 2015, 2016). The impact
of white mold on the dry bean industry in the Northwestern United States
alone has been estimated at a loss of 140 kg/ha with just 10\% disease
incidence (Ramasubramaniam et al., 2008).

Currently, there are no commercially available resistant cultivars of
dry bean (Otto-Hanson et al., 2011). Organized breeding efforts have
used a common-garden approach with white mold screening nurseries in dry
bean production areas across the United States with additional sites in
Australia, France, and Mexico (Steadman et al., 2003, 2004, 2005, 2006;
Otto-Hanson \& Steadman, 2007, 2008; McCoy \& Steadman, 2009). These
white mold screening nurseries use no chemical or cultural treatments
against \emph{S. sclerotiorum} and employ standardized protocols for
screening new cultivars for resistance to white mold (Steadman et al.,
2003; Otto-Hanson et al., 2011). These protocols included three
established cultivars used for comparison in the trials: Beryl (great
northern bean, susceptible), Bunsi (a.k.a. Ex Rico, navy bean, low
susceptibility), and G122 (cranberry bean, partial resistance) (Tu \&
Beversdorf, 1982; Steadman et al., 2005; Otto-Hanson et al., 2011). It
was previously shown that aggressiveness (the severity of disease
symptoms on the host) is significantly different across white mold
screening nursery sites in separate geographic regions (Otto-Hanson et
al., 2011). The genetic structure and mode of reproduction in these
populations, however, is currently unknown.

Understanding genetic relationships and reproduction behavior of
\emph{S. sclerotiorum} populations is beneficial for breeders seeking to
develop new resistant cultivars for worldwide deployment (Milgroom,
1996; McDonald \& Linde, 2002). In particular, genetically diverse
populations with high rates of sexual reproduction are more likely to
overcome host resistance. Most populations of \emph{S. sclerotiorum} are
predominantly clonal with low genetic diversity and have a large degree
of population fragmentation (Kohli et al., 1995; Cubeta et al., 1997;
Kohli \& Kohn, 1998; Carbone \& Kohn, 2001; Ekins et al., 2011;
Attanayake et al., 2012). Some studies, however have found populations
that show signatures of sexual reproduction (Atallah et al., 2004;
Sexton \& Howlett, 2004; Attanayake et al., 2013; Aldrich-Wolfe et al.,
2015).

Nearly all population genetic studies of \emph{S. sclerotiorum} employ a
macroscopic assay to determine mycelial compatibility, the ability for
fungal hyphae from different colonies to appear to grow together without
forming a barrier of dead cells between them (known as a barrage line,
Fig. \ref{mcg-fig}B) (Leslie, 1993; Sirjusingh \& Kohn, 2001). Mycelial
compatibility has been used as a proxy for vegetative compatibility, a
fungal trait controlled by several independent genes controlling the
ability for two hyphae to fuse and grow as a single unit (Fig.
\ref{mcg-fig}A) (Leslie, 1993; Schafer \& Kohn, 2006). Because of the
genetic connection to vegetative compatibility, two isolates that are
mycelially compatible were considered clones (Leslie, 1993); but
correlation with genetic markers, such as microsatellites, have shown
inconsistent results (Ford et al., 1995; Micali \& Smith, 2003; Jo et
al., 2008; Attanayake et al., 2012; Papaioannou \& Typas, 2014; Lehner
et al., 2017). Thus, the relationship between mycelial compatibility
groups and clonal genotypes remains unclear.

In the present study, we analyze and characterize the genetic and
phenotypic diversity of 366 \emph{S. sclerotiorum} isolates collected
between 2003 and 2012 from dry bean cultivars among different geographic
locations in the Australia, France, Mexico, and the United States. We
wanted to know if the \emph{S. sclerotiorum} populations from white mold
screening nurseries were representative of the producer fields within
the same region. As these nurseries were not treated with any chemical
or cultural control of white mold, we hypothesized that these nurseries
would represent the natural population of \emph{S. sclerotiorum}.
Furthermore, we wanted to investigate the potential effect of cultivar
on genetic diversity of the pathogen by assessing three dry bean
cultivars with different levels of resistance, Beryl (great northern
bean, susceptible), Bunsi (navy bean, low susceptibility), and G122
(cranberry bean, partial resistance) (Otto-Hanson et al., 2011). We
additionally wanted to determine categorical or phenotypic variables
that best predicted genetic structure and if there was correlation
between multilocus haplotype and mycelial compatibility group. Knowing
what variables predict genetic structure can help direct breeding
efforts. By investigating these aims, we will effectively describe the
population structure of \emph{S. sclerotiorum} in the USA and make
available our database of isolates for use in future dry bean breeding
efforts.

\section*{Materials and Methods}\label{materials-and-methods}
\addcontentsline{toc}{section}{Materials and Methods}

\subsection*{Isolate collection}\label{isolate-collection}
\addcontentsline{toc}{subsection}{Isolate collection}

Several (156) of the isolates used for this study were collected as
reported in previous studies using the same methods (Otto-Hanson et al.,
2011). Broadly, isolates were collected from two sources: white mold
screening nurseries (wmn) or producer fields. White mold screening
nurseries were 5m x 10m in size and maintained without application of
fungicides to observe natural incidence of white mold. The early nursery
plots were incorporated with a basal dressing of N:P:K = 1:3:2 and side
dressing of 0:3:2 during the growing season (Steadman et al., 2003).

Sampling was carried out by collecting sclerotia from diseased tissue in
zig-zag transects across field plots. Because sampling depended on
disease incidence, the number of samples isolated varied from year to
year. Although the nursery locations were the same over sampling years,
sampling plots within a location varied for sampling years.

Sclerotia of \emph{S. sclerotiorum} were collected over several years
from grower fields and/or wmn in 11 states of the Australia, France,
Mexico, and the United States (Table \ref{tab:isolate-table}). After
collection, sclerotia were stored in Petri plates lined with filter
paper, then stored at \(20~^{\circ}\)F or -\(4~^{\circ}\)C. Sclerotia
were surface-sterilized with 50\% Clorox bleach (at least 6\% NaOCl, The
Clorox Company, Oakland, CA) solution for 3 min, and double rinsed with
ddH\(_2\)O for 3 min. The sterilized sclerotia were then placed on water
agar plates (16g of Bacto agar per liter of ddH\(_2\)O, BD Diagnostic
Systems, Sparks, MD), with four to five sclerotia of each isolate
separated on each plate and stored on the counter top at room
temperature for 5 to 6 days. An 8-mm plug from a 5- or 6-day-old culture
was transferred from the advancing margin of the mycelia onto a plate of
Difco potato dextrose agar (PDA at 39 g/liter of ddH\(_2\)O)
(Otto-Hanson et al., 2011). In combination with the 156 isolates
described previously, we collected 210 isolates for a total of 366
isolates (Otto-Hanson et al., 2011).

\subsection*{Mycelial compatibility}\label{mycelial-compatibility}
\addcontentsline{toc}{subsection}{Mycelial compatibility}

MCG was determined as described previously through co-culturing pairs of
2-day-old isolates 2.5 cm apart on Diana Sermons (DS) Medium (Fig.
\ref{mcg-fig}) (Cubeta et al., 2001). Incompatibility of different MCGs
resulted in formation of a barrage line accompanied by formation of
sclerotia on either side of the barrage line, indicating the limits of
mycelial growth (Kohn et al., 1990; Leslie, 1993; Otto-Hanson et al.,
2011). Isolates were compared in a pairwise manner for each site and
then representatives among sites were compared to determine mycelial
compatibility groups by scoring compatible and incompatible interactions
(Otto-Hanson et al., 2011). No MCGs were compatible with any other MCG.

\subsection*{Aggressiveness}\label{aggressiveness}
\addcontentsline{toc}{subsection}{Aggressiveness}

Aggressiveness of each isolate was assessed using a straw test as
described in Otto-Hanson et al. (2011) that rated necrotic lesion size
(Petzoldt \& Dickson, 1996; Teran et al., 2006). Briefly, the straw test
uses 21-day-old G122 plants as the host in a greenhouse setting. Clear
drinking straws cut to 2.5 cm and heat sealed were used to place two
mycelial plugs of inoculum on the host plant after removing plant growth
beyond 2.5 cm above the fourth node. Measurements of the necrotic lesion
were taken 8 days later using the Modified Petzoldt and Dickson scale of
1--9, where 1 is no disease and 9 is plant death (Petzoldt \& Dickson,
1996; Teran et al., 2006).

\subsection*{Microsatellite genotyping}\label{microsatellite-genotyping}
\addcontentsline{toc}{subsection}{Microsatellite genotyping}

Prior to DNA extraction, isolates were grown on PDA and plugs were
subsequently transferred to Potato Dextrose Broth (PDB) where they were
grown until there was significant mycelial growth, but before the
mycelial mat became solidified (4--5 days). Each mycelial mat was
collected in a filtered Büchner funnel, agar plugs removed, lyophilized
and pulverized manually in Whirl-pak® HDPE sampling bags (Sigma-Aldrich,
St.~Louis, MO). Lyophilized mycelia was then stored in microcentrifuge
tubes at -\(20~^{\circ}\)C until needed for DNA extraction. DNA from
25mg of pulverized mycelia was purified using a phenol-chloroform
extraction method followed by alcohol precipitation and evaporation,
suspending the DNA in 200\(\mu\)l TE (Sambrook et al., 1989). Suspended
DNA was stored at \(4~^{\circ}\)C until genotyping.

We genotyped each \emph{S. sclerotiorum} isolate using 16 microsatellite
primer pairs developed previously (Sirjusingh \& Kohn, 2001). PCR was
carried out as described previously, using primers labeled with FAM
fluorophore. Resulting amplicons were first resolved in a 1.5\% agarose
gel stained with ethidium bromide to ensure product was within the
expected size range prior to capillary electrophoresis. Capillary
electrophoresis (fragment analysis) of amplicons, with size standard
GeneScan™ 500 LIZ®, was performed using an ABI 3730 genetic analyzer
(Life Technologies Corporation, Carlsbad, CA) at the Michigan State
University Genomic Sequencing Center (East Lansing, MI). Alleles were
scored using PeakScanner version 1.0 (Life Technologies Corporation,
Carlsbad, CA) and recorded manually in a spreadsheet.

\subsection*{Data processing and
analysis}\label{data-processing-and-analysis}
\addcontentsline{toc}{subsection}{Data processing and analysis}

All data processing and analyses were performed in a Rocker ``verse''
project container running R version 3.4.2 (Boettiger \& Eddelbuettel,
2017; R Core Team, 2017) and are openly available and reproducible at
\url{https://github.com/everhartlab/sclerotinia-366/}. Of the 16
microsatellite loci genotyped, five included compound repeats, which
made it challenging to accurately/confidently bin alleles into fragment
sizes expected for each locus based on the described repeat motif. Loci
with compound repeats were removed for the reported statistics. To
ensure the integrity of the results we additionally processed these loci
and included them in concurrent analyses. We assessed the power of our
11 markers by generating a genotype accumulation curve in the R package
\emph{poppr} version 2.5.0, looking for evidence of saturation, which
would indicate that loci were sufficiently sampled to adequately
represent the full set of haplotypes (Arnaud-Hanod et al., 2007; Kamvar
et al., 2015). To avoid including isolates potentially collected from
the same plant (which increases the probability of collecting sclerotia
from the same point of infection more than once), data were
clone-corrected on a hierarchy of Region/Source/Host/Year---meaning that
duplicated genotypes were reduced to a single observation when they were
collected in the same year from the same host cultivar located in the
same source field (wmn or producer)---for subsequent analysis. We
assessed haplotype diversity by calculating Stoddart and Taylor's index
(\emph{G}) (Stoddart \& Taylor, 1988), Shannon's index (\emph{H})
(Shannon, 1948), Simpson's index (\(\lambda\)) (Simpson, 1949), evenness
(\(E_5\)), and the expected number of multilocus haplotypes
(\emph{eMLH}) (Hurlbert, 1971; Heck et al., 1975; Pielou, 1975; Grünwald
et al., 2003). To assess the potential for random mating, we tested for
linkage disequilibrium with the index of association, \(I_A\) and its
standardized version, \(\bar{r}_d\) using 999 permutations (Brown et
al., 1980; Smith et al., 1993; Agapow \& Burt, 2001). Both haplotype
diversity and linkage disequilibrium were calculated in \emph{poppr}
(Kamvar et al., 2014).

\subsection*{Assessing importance of
variables}\label{assessing-importance-of-variables}
\addcontentsline{toc}{subsection}{Assessing importance of variables}

\subsubsection*{Distance-based redundancy
analysis}\label{distance-based-redundancy-analysis}
\addcontentsline{toc}{subsubsection}{Distance-based redundancy analysis}

A distance-based redundancy analysis (dbRDA) (Legendre \& Anderson,
1999) was performed with the function \texttt{capscale()} in the
\emph{vegan} package version 2.4.4 (Oksanen et al., 2017). This method
uses constrained ordinations on a distance matrix representing the
response variable to delineate relative contribution of any number of
independent explanatory variables. We used this method to delineate the
phenotypic (Aggressiveness, Mycelial Compatibility Group (MCG)),
geographic (Region, Host, Location), and temporal (Year) components in
predicting genetic composition of the populations. The distance matrix
we used as the response variable was generated using Bruvo's genetic
distance from clone-corrected data (procedure described above) as
implemented in \emph{poppr}, which employed a stepwise mutation model
for microsatellite data (Bruvo et al., 2004; Kamvar et al., 2014).
Because aggressiveness measures differed between isolates that were
reduced to a single observation during clone-correction, aggressiveness
was first averaged across clone-corrected isolates. To identify
explanatory variable(s) correlated with genetic variation, a
forward-backward selection process was applied with the vegan function
\texttt{ordistep()}. An analysis of variance (ANOVA) was then performed
to test for significance of the reduced model and marginal effects using
999 permutations. The \texttt{varpart()} function of vegan was used to
determine variation partitioning of explanatory variables.

\subsubsection*{Aggressiveness
assessment}\label{aggressiveness-assessment}
\addcontentsline{toc}{subsubsection}{Aggressiveness assessment}

We used ANOVA to assess if aggressiveness (determined via straw test on
a scale of 1--9 as described above) was significantly different with
respect to Region, MCG, or multilocus haplotype (MLH). To minimize
complications due to small sample sizes, we chose the top 10 MCGs,
representing 56.5\% of the isolates collected, the 10 most abundant MLHs
representing 26.7\% of the isolates, and populations with more than five
isolates. If ANOVA results were significantly different at \(\alpha\) =
0.05, pairwise differences were assessed using Tukey's HSD test
(\(\alpha\) = 0.05) using the \texttt{HSD.test()} function in the
package \emph{agricolae} version 1.2.8 (Mendiburu \& Simon, 2015).

\subsubsection*{Correlating multilocus haplotypes with mycelial
compatibility
groups}\label{correlating-multilocus-haplotypes-with-mycelial-compatibility-groups}
\addcontentsline{toc}{subsubsection}{Correlating multilocus haplotypes
with mycelial compatibility groups}

We wanted to assess if there was correlation between MLHs and MCGs. This
was performed using a network-based approach where both MLHs and MCGs
were considered nodes and the number of isolates in which they were
found together was the strength of the connection between an MLH and and
MCG node. The network-based approach allowed us to assess the
associations between MLHs and MCGs. To construct the network, a
contingency table was created with MLHs and MCGs and converted to a
directed and weighted edgelist where each edge represented a connection
from an MCG to an MLH, weighted by the number of samples shared in the
connection. This was then converted to a bipartite graph where top nodes
represented MLHs and bottom nodes represented MCGs. To identify clusters
of MLHs and MCGs within the network, we used the cluster walktrap
community detection algorithm as implemented in the
\texttt{cluster\_walktrap()} function in \emph{igraph} version 1.1.2
(Csardi \& Nepusz, 2006; Pons \& Latapy, 2006). This algorithm attempts
to define clusters of nodes by starting at a random node and performing
short, random ``walks'' along the edges between nodes, assuming that
these walks would stay within clusters. For this analysis, we set the
number of steps within a walk to four and allowed the algorithm to use
the edge weights in determining the path. All of the resulting
communities that had fewer than 10 members were then consolidated into
one. Community definitions were used to assess the average genetic
distance (as defined by Bruvo's distance) within members of the
community (Bruvo et al., 2004).

\subsection*{Genetic diversity}\label{genetic-diversity}
\addcontentsline{toc}{subsection}{Genetic diversity}

\subsubsection*{Population
differentiation}\label{population-differentiation}
\addcontentsline{toc}{subsubsection}{Population differentiation}

We used analysis of molecular variance (AMOVA) with Bruvo's genetic
distance in \emph{poppr} to test for differentiation between populations
in wmn and producer fields from the same region and collected in the
same year (Excoffier et al., 1992; Bruvo et al., 2004; Kamvar et al.,
2014). To identify Regions with greater differentiation, we used
discriminant analysis of principal components (DAPC) as implemented in
\emph{adegenet} version 2.1.0, assessing the per-sample posterior group
assignment probability (Jombart, 2008). This method decomposes the
genetic data into principal components, and then uses a subset of these
as the inputs for discriminant analysis, which attempts to minimize
within-group variation and maximize among-group variation (Jombart et
al., 2010). To avoid over-fitting data, the optimal number of principal
components was selected by using the \emph{adegenet} function
\texttt{xvalDapc()}. This function implements a cross-validation
procedure to iterate over an increasing number of principal components
on a subset (90\%) of the data, trying to find the minimum number of
principal components that maximizes the rate of successful group
reassignment. To assess if cultivar had an influence on genetic
diversity between wmn, we first subset the clone-corrected data to
contain only samples from wmn and from the cultivars Beryl, Bunsi, and
G122 and tested differentiation using AMOVA and DAPC as described above.
We additionally assessed population stability over time by calculating
DAPC over the combined groups of Region and Year as described above.

\subsubsection*{Analysis of shared multilocus
haplotypes}\label{analysis-of-shared-multilocus-haplotypes}
\addcontentsline{toc}{subsubsection}{Analysis of shared multilocus
haplotypes}

We wanted to evaluate patterns of connectivity between shared multilocus
haplotypes across geographic regions. We first tabulated the multilocus
haplotypes shared between at least two populations (defined as states or
countries) with the \emph{poppr} function \texttt{mlg.crosspop()}
(Kamvar et al., 2014). From these data, we constructed a graph with
populations as nodes and shared haplotypes as edges (connections)
between nodes using the R packages \emph{igraph} (Csardi \& Nepusz,
2006), \emph{dplyr} version 0.7.4 (Wickham et al., 2017), and
\emph{purrr} version 0.2.4 (Henry \& Wickham, 2017). Each node was
weighted by the fraction of shared MLHs. Each edge represented a single
MLH, but because a single MLH could be present in more than one
population, that MLH would have a number of edges equivalent to the
total number of possible connections, calculated as
(\emph{n}*(\emph{n}-1))/2 edges where n represents the number of
populations crossed. Edges were weighted by \(1-P_{sex}\), where
\(P_{sex}\) is the probability of encountering the same haplotype via
two independent meiotic events (Parks \& Werth, 1993; Arnaud-Hanod et
al., 2007). This weighting scheme would thus strengthen the connection
of edges that represented genotypes with a low probability of being
produced via sexual reproduction. We then identified communities (among
the Regions) in the graph using the \texttt{cluster\_optimal()} function
from \emph{igraph} (Csardi \& Nepusz, 2006). The graph was plotted using
the R packages \emph{ggplot2} version 2.2.1 (Wickham, 2009) and
\emph{ggraph} 1.0.0 (Pedersen, 2017). To ensure that we captured the
same community signal, we additionally performed this analysis including
the five polymorphic markers described above.

\section*{Results}\label{results}
\addcontentsline{toc}{section}{Results}

A total of 366 isolates were collected from 2003 to 2012 (except 2006
and 2011) from diseased dry bean plants in 11 states in the United
States as well as Australia, France, and Mexico (Table
\ref{tab:isolate-table}). With the 11 loci used in the analyses (Table
\ref{tab:locus-stats}), we observed a total of 165 MLHs (215 with 16
loci). These 11 loci are located on 7 chromosomes in the \emph{S.
sclerotiorum} genome with a minimum distance of 55Kbp between two loci
on the same chromosome. Over 50\% of the isolates came from four states,
MI (62), ND (60), WA (59), NE (47). Four regions had fewer than 10
isolates, Australia (6), WI (2), NY (1), ID (1). We observed 87 MCGs,
the most abundant of which (`MCG 5') was represented by 73 isolates over
37 MLHs (Fig. \ref{biggraph}A,C).

The number of observed alleles per locus ranged from two to 10 with an
average of 6.27 (Table \ref{tab:locus-stats}). Locus 20-3, which
contained only 2 alleles, showed low values of both \emph{h} (0.0533)
and evenness (0.42), indicating that there was one dominant allele
present. Analysis of the haplotype accumulation curve showed no clear
plateau for 11 or 16 loci (See section on `Loading Data and Setting
Strata' in the MLG-distribution.md\footnote{Direct link:
  \url{https://github.com/everhartlab/sclerotinia-366/blob/master/results/MLG-distribution.md\#loading-data-and-setting-strata}}
file in the supplemental files (Kamvar et al., 2017)), indicating that
we would likely obtain more multilocus haplotypes if we were to genotype
more loci.

After clone-correction on the hierarchy of Region/Source/Host/Year, a
total of 48 isolates were removed from the data set, resulting in 318
isolates representing 165 MLHs that were used in subsequent analyses
(Table \ref{tab:iatab}). The results showed that, in terms of genotypic
diversity (\emph{H}, \emph{G}, and \(\lambda\)), WA was the most diverse
population with both \emph{G} (54.3) and \(e^H\) (55.3) being close to
the observed number of MLHs (56). This indicated that there are few
duplicated genotypes in WA (Table \ref{tab:iatab}). A more useful metric
to compare populations, however, is \(E_5\), which scales from 0 to 1,
where 1 indicates all unique genotypes (Grünwald et al., 2003).
Evaluating by \(E_5\) shows that both MI and NE exhibit lower than
average values, indicating that there are over-represented genotypes in
the popualtions (table \ref{tab:iatab}). When we look at Mexico, we
observed that it had relatively high values of \(E_5\) and genotypic
diversity, but low richness, as measured by \emph{eMLG}. Moreover,
Mexico had the lowest value for \emph{h}, which is a measure of allelic
diversity. Nearly all populations showed evidence of linkage (Table
\ref{tab:iatab}), which serves as evidence for clonal reproduction or
other forms of non-random mating. The only exceptions were CA (\emph{P}
= 0.043) and Australia (\emph{P} = 0.052). Both of these populations
showed only moderate significance with \(\bar{r}_d\) values of 0.03 and
0.12, respectively.

\begin{longtable}[]{@{}rrrlrl@{}}
\caption{\label{tab:locus-stats} Allelic diversity on full data set at loci
used in this study. \emph{h} = Nei's Gene Diversity (Nei, 1978). Average
\emph{h} = 0.583, average Evenness = 0.693, average no. alleles =
6.27}\tabularnewline
\toprule
\begin{minipage}[b]{0.08\columnwidth}\raggedleft\strut
Locus\strut
\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\raggedleft\strut
Range\strut
\end{minipage} & \begin{minipage}[b]{0.29\columnwidth}\raggedleft\strut
Repeat Motif\strut
\end{minipage} & \begin{minipage}[b]{0.14\columnwidth}\raggedright\strut
No. alleles\strut
\end{minipage} & \begin{minipage}[b]{0.06\columnwidth}\raggedleft\strut
\emph{h}\strut
\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\raggedright\strut
Evenness\strut
\end{minipage}\tabularnewline
\midrule
\endfirsthead
\toprule
\begin{minipage}[b]{0.08\columnwidth}\raggedleft\strut
Locus\strut
\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\raggedleft\strut
Range\strut
\end{minipage} & \begin{minipage}[b]{0.29\columnwidth}\raggedleft\strut
Repeat Motif\strut
\end{minipage} & \begin{minipage}[b]{0.14\columnwidth}\raggedright\strut
No. alleles\strut
\end{minipage} & \begin{minipage}[b]{0.06\columnwidth}\raggedleft\strut
\emph{h}\strut
\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\raggedright\strut
Evenness\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
5-2\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedleft\strut
318--324\strut
\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedleft\strut
(GT)\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
4\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
0.45\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
0.62\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
6-2\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedleft\strut
483--495\strut
\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedleft\strut
(TTTTTC)(TTTTTG)(TTTTTC)\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
3\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
0.64\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
0.95\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
7-2\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedleft\strut
158--174\strut
\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedleft\strut
(GA)\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
7\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
0.73\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
0.76\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
8-3\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedleft\strut
244--270\strut
\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedleft\strut
(CA)\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
7\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
0.74\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
0.79\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
9-2\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedleft\strut
360--382\strut
\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedleft\strut
(CA)(CT)\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
9\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
0.35\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
0.41\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
12-2\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedleft\strut
214--222\strut
\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedleft\strut
(CA)\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
5\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
0.58\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
0.78\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
17-3\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedleft\strut
342--363\strut
\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedleft\strut
(TTA)\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
7\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
0.55\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
0.53\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
20-3\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedleft\strut
280--282\strut
\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedleft\strut
(GT)GG(GT)\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
2\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
0.05\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
0.42\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
55-4\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedleft\strut
153--216\strut
\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedleft\strut
(TACA)\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
10\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
0.72\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
0.66\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
110-4\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedleft\strut
370--386\strut
\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedleft\strut
(TATG)\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
5\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
0.76\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
0.91\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
114-4\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedleft\strut
339--416\strut
\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedleft\strut
(TAGA)\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
10\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
0.83\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
0.80\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

\newpage

\begin{longtable}[]{@{}crrrrllrl@{}}
\caption{\label{tab:iatab} Genotypic diversity and Linkage Disequilibrium
summary for geographic populations arranged by abundance after
clone-correction by a hierarchy of Region/Source/Host/Year. Pop =
Population, N = number of individuals (number of MLH in parentheses),
eMLH = expected number of MLHs based on rarefaction at 10 individuals
(standard error in parentheses), H = Shannon-Weiner Index, G = Stoddardt
and Taylor's Index, \(\lambda\) = Simpson's Index, \emph{h} = Nei's 1978
gene diversity, \(E_5\) = Evenness, \(\bar{r}_d\) = standardized index
of association. An asterix indicates a significant value of
\(\bar{r}_d\) after 999 permutations, \(P \leq 0.001\).}\tabularnewline
\toprule
\begin{minipage}[b]{0.12\columnwidth}\centering\strut
Pop\strut
\end{minipage} & \begin{minipage}[b]{0.08\columnwidth}\raggedleft\strut
N\strut
\end{minipage} & \begin{minipage}[b]{0.13\columnwidth}\raggedleft\strut
eMLH\strut
\end{minipage} & \begin{minipage}[b]{0.04\columnwidth}\raggedleft\strut
H\strut
\end{minipage} & \begin{minipage}[b]{0.05\columnwidth}\raggedleft\strut
G\strut
\end{minipage} & \begin{minipage}[b]{0.11\columnwidth}\raggedright\strut
\(\lambda\)\strut
\end{minipage} & \begin{minipage}[b]{0.06\columnwidth}\raggedright\strut
\(E_5\)\strut
\end{minipage} & \begin{minipage}[b]{0.06\columnwidth}\raggedleft\strut
\emph{h}\strut
\end{minipage} & \begin{minipage}[b]{0.07\columnwidth}\raggedright\strut
\(\bar{ r}_d\)\strut
\end{minipage}\tabularnewline
\midrule
\endfirsthead
\toprule
\begin{minipage}[b]{0.12\columnwidth}\centering\strut
Pop\strut
\end{minipage} & \begin{minipage}[b]{0.08\columnwidth}\raggedleft\strut
N\strut
\end{minipage} & \begin{minipage}[b]{0.13\columnwidth}\raggedleft\strut
eMLH\strut
\end{minipage} & \begin{minipage}[b]{0.04\columnwidth}\raggedleft\strut
H\strut
\end{minipage} & \begin{minipage}[b]{0.05\columnwidth}\raggedleft\strut
G\strut
\end{minipage} & \begin{minipage}[b]{0.11\columnwidth}\raggedright\strut
\(\lambda\)\strut
\end{minipage} & \begin{minipage}[b]{0.06\columnwidth}\raggedright\strut
\(E_5\)\strut
\end{minipage} & \begin{minipage}[b]{0.06\columnwidth}\raggedleft\strut
\emph{h}\strut

>>>>

\input{paper}
\newpage
\input{supplementary}
\end{document}
